---
title: "Thesis Data Analysis"
author: "Adam W. Gelman"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  - \usepackage{changepage}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, tidy=TRUE, include=TRUE)
library(rempsyc)
library(broom)
require(tinytex)
require(xtable)
require(olsrr)
require(mosaic)
require(Sleuth2)
require(leaps)
require(MASS)
require(effectsize)
require(flextable)
require(report)
```

```{r Data_Import, echo=FALSE}
library(readxl)
Master_Spreadsheet <- read_excel("Master_Spreadsheet.xlsx", na = ".")
```

# Single-Variable t-Test p-Values by Response Variable

\begin{adjustwidth}{2cm}{2cm}

The following table compares the performance of plants grown in 
Perlite with those grown in Regolith without considering microbial
inoculation. P-values greater than 0.05 indicate a high probability 
that the observed difference occurred due to random chance.

\end{adjustwidth}

```{r Substrate-t-Tests, echo=FALSE, results='asis', fig.width=6.5}

table1=xtable(nice_t_test(
  data = Master_Spreadsheet,
  response = c("Grade_1_Count","Shoot_Fresh_Weight","Shoot_Dry_Weight"),
  group = "Substrate"
))
print(table1, comment=FALSE)

```

\begin{adjustwidth}{2cm}{2cm}

The following table compares the performance of plants grown 
with and without microbial inoculation without considering the
substrate in which the plants were grown. P-values greater than 
0.05 indicate a greater than 5% probability that the observed 
difference occurred due to random chance.

\end{adjustwidth}

```{r Inoculation-t-Tests, echo=FALSE, results='asis', fig.width=6.5}
table3=xtable(nice_t_test(
  data = Master_Spreadsheet,
  response = c("Grade_1_Count","Shoot_Fresh_Weight","Shoot_Dry_Weight"),
  group = "Inoculation"
))

print(table3, comment=F)
```

# Group-by-Group t-Test p-Values

\begin{adjustwidth}{2cm}{2cm}

The following table provides p-values for comparisons between 
each treatment group and each of the three other treatment groups.

\end{adjustwidth}

```{r t-Tables, echo=FALSE, results='asis', fig.width=6.5}

# Subset for Substrate = "R" and Inoculation = "M"
subset_RM <- subset(Master_Spreadsheet, Substrate == "R" & Inoculation == "M")

# Subset for Substrate = "R" and Inoculation = "NM"
subset_R <- subset(Master_Spreadsheet, Substrate == "R" & Inoculation == "NM")

# Subset for Substrate = "P" and Inoculation = "M"
subset_PM <- subset(Master_Spreadsheet, Substrate == "P" & Inoculation == "M")

# Subset for Substrate = "P" and Inoculation = "NM"
subset_P <- subset(Master_Spreadsheet, Substrate == "P" & Inoculation == "NM")

RvP_G1Count = t.test(Master_Spreadsheet$Grade_1_Count~Master_Spreadsheet$Substrate)
MvNM_G1Count= t.test(Master_Spreadsheet$Grade_1_Count~Master_Spreadsheet$Inoculation)
RvP_G1Count = t.test(Master_Spreadsheet$Shoot_Fresh_Weight~Master_Spreadsheet$Substrate)
MvNM_G1Count= t.test(Master_Spreadsheet$Shoot_Fresh_Weight~Master_Spreadsheet$Inoculation)
RvP_G1Count = t.test(Master_Spreadsheet$Shoot_Dry_Weight~Master_Spreadsheet$Substrate)
MvNM_G1Count= t.test(Master_Spreadsheet$Shoot_Dry_Weight~Master_Spreadsheet$Inoculation)



RMvR_G1Count = tidy(t.test(subset_RM$Grade_1_Count,subset_R$Grade_1_Count))
RMvP_G1Count = tidy(t.test(subset_RM$Grade_1_Count,subset_P$Grade_1_Count))
RMvPM_G1Count = tidy(t.test(subset_RM$Grade_1_Count,subset_PM$Grade_1_Count))
RvP_G1Count = tidy(t.test(subset_R$Grade_1_Count,subset_P$Grade_1_Count))
RvPM_G1Count = tidy(t.test(subset_R$Grade_1_Count,subset_PM$Grade_1_Count))
PvPM_G1Count = tidy(t.test(subset_P$Grade_1_Count,subset_PM$Grade_1_Count))

RMvR_Fresh_Weight = tidy(t.test(subset_RM$Shoot_Fresh_Weight,subset_R$Shoot_Fresh_Weight))
RMvP_Fresh_Weight = tidy(t.test(subset_RM$Shoot_Fresh_Weight,subset_P$Shoot_Fresh_Weight))
RMvPM_Fresh_Weight = tidy(t.test(subset_RM$Shoot_Fresh_Weight,subset_PM$Shoot_Fresh_Weight))
RvP_Fresh_Weight = tidy(t.test(subset_R$Shoot_Fresh_Weight,subset_P$Shoot_Fresh_Weight))
RvPM_Fresh_Weight = tidy(t.test(subset_R$Shoot_Fresh_Weight,subset_PM$Shoot_Fresh_Weight))
PvPM_Fresh_Weight = tidy(t.test(subset_P$Shoot_Fresh_Weight,subset_PM$Shoot_Fresh_Weight))

RMvR_Dry_Weight = tidy(t.test(subset_RM$Shoot_Dry_Weight,subset_R$Shoot_Dry_Weight))
RMvP_Dry_Weight = tidy(t.test(subset_RM$Shoot_Dry_Weight,subset_P$Shoot_Dry_Weight))
RMvPM_Dry_Weight = tidy(t.test(subset_RM$Shoot_Dry_Weight,subset_PM$Shoot_Dry_Weight))
RvP_Dry_Weight = tidy(t.test(subset_R$Shoot_Dry_Weight,subset_P$Shoot_Dry_Weight))
RvPM_Dry_Weight = tidy(t.test(subset_R$Shoot_Dry_Weight,subset_PM$Shoot_Dry_Weight))
PvPM_Dry_Weight = tidy(t.test(subset_P$Shoot_Dry_Weight,subset_PM$Shoot_Dry_Weight))

G1Count_P_Values = c(
  RMvR_G1Count$p.value,
  RMvP_G1Count$p.value,
  RMvPM_G1Count$p.value,
  RvP_G1Count$p.value,
  RvPM_G1Count$p.value,
  PvPM_G1Count$p.value)

Fresh_Weight_P_Values = c(
  RMvR_Fresh_Weight$p.value,
  RMvP_Fresh_Weight$p.value,
  RMvPM_Fresh_Weight$p.value,
  RvP_Fresh_Weight$p.value,
  RvPM_Fresh_Weight$p.value,
  PvPM_Fresh_Weight$p.value)

Dry_Weight_P_Values = c(
  RMvR_Dry_Weight$p.value,
  RMvP_Dry_Weight$p.value,
  RMvPM_Dry_Weight$p.value,
  RvP_Dry_Weight$p.value,
  RvPM_Dry_Weight$p.value,
  PvPM_Dry_Weight$p.value)

P_Table_Cols = c("RM Vs. R","RM Vs. P","RM Vs. PM","R Vs. P","R Vs. PM","P Vs. PM")
P_Table_Rows = c("Count of Green Fruit","Fresh Weight","Dry Weight")

table=data.frame(Comparison=P_Table_Cols, Grade_1_Count=G1Count_P_Values, Fresh_Weight=Fresh_Weight_P_Values, Dry_Weight=Dry_Weight_P_Values)

table2=xtable(table)
print(table2, comment=FALSE)

```

# Boxplots

\begin{adjustwidth}{2cm}{2cm}

The following boxplots allow visual comparison of each of the 
four treatment groups as analyzed using t-tests in the previous 
step. These boxplots are marked with letters which show 
statistically significant differences; boxplots with the same 
letter represent groups that are NOT significantly different 
from one another.

The black line represents the median value of the group.
The black dot represents the mean of each group. The grey 
box represents the middle 50% of observations. The whiskers 
represent the furthest edge cases within 1.5 times the 
interquartile range from the 25th and 75th percentiles. 
Circles beyond the whiskers represent possible outliers.

\end{adjustwidth}

## Count of Green Fruit

```{r Grade_One_Count_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        Grade_1_Count ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Count\nof\nGreen\nFruit\n(Grade 1)", side = 2, line = 4, at = mean(range(Master_Spreadsheet$Grade_1_Count)), las = 1, adj=0.5)

means <- aggregate(Grade_1_Count ~ Inoculation + Substrate, data = Master_Spreadsheet, FUN = mean)$Grade_1_Count

points(1:4, means, col = "black", pch=16)

text(1:4, means + 3,  # `means + 2` adjusts the position above the mean bars
     labels = c('a', 'a', 'b', 'ab'), 
     cex = 0.8)

```

## Count of Red Fruit

```{r Grade_Five_Six_Count_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        Grade_5_6_Count ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Count\nof\nRed\nFruit\n(Grades 5 & 6)", side = 2, line = 4, at = mean(range(Master_Spreadsheet$Grade_5_6_Count)), las = 1, adj=0.5)

means <- aggregate(Grade_5_6_Count ~ Inoculation + Substrate, data = Master_Spreadsheet, FUN = mean)$Grade_5_6_Count

points(1:4, means, col = "black", pch=16)

text(1:4, means + 0.5,
     labels = c('a', 'a', 'a', 'a'), 
     cex = 0.8)
```

## Shoot Fresh Weight

```{r Shoot_Fresh_Weight_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        Shoot_Fresh_Weight ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Shoot\nFresh\nWeight (g)", side = 2, line = 4, at = mean(range(Master_Spreadsheet$Shoot_Fresh_Weight)), las = 1, adj=0.5)

means <- aggregate(Shoot_Fresh_Weight ~ Inoculation + Substrate, data = Master_Spreadsheet, FUN = mean)$Shoot_Fresh_Weight

points(1:4, means, col = "black", pch=16)

text(1:4, means -6,
     labels = c('a', 'a', 'b', 'b'), 
     cex = 0.8)
```

## Shoot Dry Weight

```{r Shoot_Dry_Weight_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        Shoot_Dry_Weight ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Shoot\nDry\nWeight (g)", side = 2, line = 4, at = mean(range(Master_Spreadsheet$Shoot_Dry_Weight)), las = 1, adj=0.5)

means <- aggregate(Shoot_Dry_Weight ~ Inoculation + Substrate,
                   data = Master_Spreadsheet,
                   FUN = mean)$Shoot_Dry_Weight

points(1:4, means, col = "black", pch=16)

text(1:4, means - 1,
     labels = c('a', 'a', 'b', 'b'), 
     cex = 0.8)
```

## Red Fruit Weight

```{r Red_Fruit_Weight_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        Grade_5_6_Weight ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Red\nFruit\n(Grades 5\nand 6)\nFresh\nWeight (g)", side = 2, line = 4, at = mean(range(Master_Spreadsheet$Grade_5_6_Weight)), las = 1, adj=0.5)

means <- aggregate(Grade_5_6_Weight ~ Inoculation + Substrate,
                   data = Master_Spreadsheet,
                   FUN = mean)$Grade_5_6_Weight

points(1:4, means, col = "black", pch=16)

text(1:4, means - 12,
     labels = c('a', 'a', 'a', 'a'), 
     cex = 0.8)
```

## Total Edible Biomass

```{r All_Grades_Weight_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        All_Grades_Weight ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Total\nEdible\nBiomass\n(g)", side = 2, line = 4, at = mean(range(Master_Spreadsheet$All_Grades_Weight)), las = 1, adj=0.5)

means <- aggregate(All_Grades_Weight ~ Inoculation + Substrate,
                   data = Master_Spreadsheet,
                   FUN = mean)$All_Grades_Weight

points(1:4, means, col = "black", pch=16)

text(1:4, means - 12,
     labels = c('a', 'a', 'a', 'a'), 
     cex = 0.8)
```

## Brix

```{r Brix_Box_Plots, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center'}
par(mar = c(5, 6, 4, 2) + 1)
boxplot(
        Average_Brix ~ Inoculation*Substrate, 
        
        data=Master_Spreadsheet, 
        xlab="Treatment", 
        ylab="", 
        names=c('Inoculated\nPerlite',
                'Non-inoculated\nPerlite',
                'Inoculated\nRegolith',
                'Non-inoculated\nRegolith'),
        cex.axis=0.63,
        las=1
        )
mtext("Average\nBrix", side = 2, line = 4, at = mean(range(Master_Spreadsheet$Average_Brix)), las = 1, adj=0.5)

means <- aggregate(Average_Brix ~ Inoculation + Substrate,
                   data = Master_Spreadsheet,
                   FUN = mean)$Average_Brix

points(1:4, means, col = "black", pch=16)

text(1:4, means - 0.05,
     labels = c('a', 'a', 'a', 'a'), 
     cex = 0.8)
```

```{r, xtable-&-R-Version-Statement, echo=FALSE, fig.height=7, fig.width=5.5, fig.align='center', results='asis'}
cat(
paste0(
  "% latex tables generated in ", 
  sub(" \\(.*$", "", R.version$version.string), 
  " by xtable ", 
  packageVersion("xtable"), 
  " package % ",
  format(Sys.time(), "%a %b %d %H:%M:%S %Y"))
)
```